### 2017-11-16 Risk analysis/Text/Layout/Layout/chapter3_results.tex
Results text pre-finished!

		To generate a set of reference results, a prior model was first generated by running a simple simulation of Monte Carlo sampling and error propagation regarding only the uncertainty of prior parameters. No likelihood functions were included in this prior model. Prior parameter uncertainties were chosen to be identical for all models, prior and posteriors.

@@

		The nature of the prior uncertainties is reflected by the visualization using information entropy (see Figure \ref{fig:ML1} (a)). High entropies are found at and near the possible locations of layer boundaries. This effect is amplified by the offset uncertainty in the footwall, where the highest information entropies are found.\\
		Regarding the different mechanisms of maximum trap volume control, it can be observed (see Figure \ref{fig:ML1} (c)) that all four relevant mechanisms occurred. The dominant factor is the anticlinal spill point (labeled "SPILL POINT"), accounting for more than 60\% of model realizations. It is followed by cross-fault leakage due to juxtaposition with formations underlying ("LEAK UNDER", ~20\%) and formations overlying ("LEAK OVER", ~13\%) the seal in the footwall. Leakage due to stratigraphical connections to seal-overlying units ("STRAT"), i.e. stratigraphical breaches of the seal, were registered to be decisive in only about 3\% of realizations. A satisfactory reliability of model algorithms and computation is indicated by the lack of unrecognized mechanisms (labeled "UNCLEAR", see Figure \ref{fig:ML1} (c)).\\
		Recoverable oil volumes were calculated for each model iteration and plotted as a probability distribution in Figure \ref{fig:M_prior} (a). It can clearly be seen that while a large portion of the distribution is found in highly positive volumes, forming a normal distribution shape around 3.8 million $m³$, there also a high probability of complete failure, i.e. a maximum trap volume of zero. This results in a bimodal characterization, with low probability of the $ROV$ to be found between zero and three million $m³$.\\
		Consequently, applying the custom loss function on this distribution results in divergent Bayes actions for the different risk-affine actors (see Figure \ref{fig:M_prior} (b)). Although all actors find their estimates to be positive, only the risk-neutral to risk-friendliest actors come close to the positive mode of the distribution. Risk-averse individuals bid on significantly lower estimates. It can also be observed, that expected losses seem to be generally lower for risk-affine actors. The different Bayes actions have apparently been torn apart by the bimodality of the $ROV$ probability distribution.

		\subsection{Posterior model I}
		\begin{table}[h]
			\centering
			\begin{tabular}{lrr} 
				\toprule
				Model I - Likelihoods: layer thicknesses (normal distributions)\\  
				\midrule 
				Formation & $\mu$ [m] & $\sigma$ [m]\\ 
				\midrule 
				Sandstone 2 & 150 & 30 \\
				Shale (Seal) & 350 & 30\\ 
				Sandstone 1 (Reservoir) & 250 & 30 \\
				\bottomrule
			\end{tabular}
			\caption{Layer thickness likelihoods determined by normal distributions as used for posterior model I.}
			\label{tab:ML1_likelihoods}
		\end{table}
		For the simulation of this first posterior model, thickness likelihoods for the three central layers were defined (see Table \ref{tab:ML1_likelihoods}).
		\begin{figure}[p!]
			\centering
			\includegraphics[width=1\textwidth]{Figures/ML1}
			\caption{....}\label{fig:ML1}
		\end{figure}

		\subsection{Posterior model II}%ML3
		-results similar to I, basically same
		- results of ML1 generally similar as for ML3:
		i.e. thicker reservoir does not mean higher volume, as we didn't implement a direct uncertainty to the shape, the spill point will stay more or less the same
		-> spill and leak point as limiting factors to max volume
		-> max volume already defined in prior (ROV max 5,500,000)

		\subsection{Posterior model III}%ML2
		\begin{figure}[p!]
			\centering
			\includegraphics[width=1\textwidth]{Figures/ML2}
			\caption{....}\label{fig:ML2}
		\end{figure}
		- see ML2: seal thickness as decisive factor
			-> if seal thin and reservoir thick see the ratio of both!), higher likelihood of surpassing SSF

>>>>

		To generate a set of reference results, a prior model was first generated by running a simple simulation of Monte Carlo sampling and error propagation regarding only the uncertainty of prior parameters. No likelihood functions were included in this prior model. Prior parameter uncertainties were chosen to be identical in all model simulations.

@@

		The nature of the prior uncertainties was visualized using information entropy (see Figure \ref{fig:ML1} (a)). High entropies are found at and near the possible locations of layer boundaries. This effect is amplified by the offset uncertainty in the footwall, where the highest information entropies are found.\\
		The occurrence of the different mechanisms of maximum trap control were traced for all simulations and labeled as follows:
		\begin{enumerate}
		\item \textbf{SPILL POINT}: Migration related to the anticlinal spill point of the trap.
		\item \textbf{LEAK UNDER}: Leakage due to juxtaposition of the reservoir with formations underlying the seal in the hanging wall.
		\item \textbf{LEAK OVER}: Leakage due to juxtaposition of the reservoir with formations overlying the seal in the hanging wall. This was defined to lead to full leakage and complete failure of the trap.
		\item \textbf{STRAT}: Leakage due to stratigraphical connections to seal-overlying units, i.e. stratigraphical breaches of the seal. Full leakage and trap failure were assumed for this mechanism as well.
		\item \textbf{UNCLEAR}: Label defined to track cases in which a mechanism was not recognized. This is deemed an indicator for failures of the algorithms or the computation, i.e. the reliability of the overall model and incorporated functions. 
		\end{enumerate}
		It is shown in Figure \ref{fig:ML1} (c) that all four relevant mechanisms occurred for the prior model. The dominant factor is the anticlinal spill point accounting for more than 60\% of model realizations. It is followed by "LEAK UNDER" ($\sim20\%$) and "LEAK OVER" ($\sim13\%$). Stratigraphical breaches of the seal were registered to be decisive in only about 3\% of iterations. A satisfactory reliability of model algorithms and computation is indicated by the lack of unrecognized mechanisms.\\
		Recoverable oil volumes were calculated for each model iteration and plotted as a probability distribution in Figure \ref{fig:M_prior} (a). It can clearly be seen that while a large portion of the distribution is found in a range of highly positive volumes, forming a normal distribution shape around 3.8 million m$^3$, there is also a high probability of complete failure, i.e. the trap volume to be zero. This results in a bimodal characterization, with low probability of the $ROV$ to be found between zero and 3 million m$^3$.\\
		Consequently, applying the custom loss function on this distribution results in divergent Bayes actions for the different risk-affine actors (see Figure \ref{fig:M_prior} (b)). Although all actors find their estimates to be positive, only the risk-neutral to risk-friendliest actors come close to the described positive mode of the distribution ($ROV > 3*10^6$ m$^3$). Risk-averse individuals bid on significantly lower estimates. It can also be observed that expected losses seem to be generally lower for risk-affine actors. The divergence of different Bayes actions presumably resulted from the bimodality of the $ROV$ probability distribution.

		\subsection{Posterior model I: Seal and reservoir likely thick}
		\begin{table}[h]
			\centering
			\begin{tabular}{lrr} 
				\toprule
				Model I - Likelihoods: layer thicknesses (normal distributions)\\  
				\midrule 
				Formation & $\mu$ [m] & $\sigma$ [m]\\ 
				\midrule 
				Sandstone 2 & 150 & 30 \\
				Shale (Seal) & 350 & 30\\ 
				Sandstone 1 (Reservoir) & 250 & 30 \\
				\bottomrule
			\end{tabular}
			\caption{Layer thickness likelihoods determined by normal distributions as used for posterior model I.}
			\label{tab:ML1_likelihoods}
		\end{table}
		\begin{figure}[p!]
			\centering
			\includegraphics[width=1\textwidth]{Figures/ML1}
			\caption{....}\label{fig:ML1}
		\end{figure}		
		For the simulation of this first posterior model, thickness likelihoods for the three central layers were included into the probabilistic model. These were defined in a way that reinforced the probabilities of the relevant layers to be significantly thick (see Table \ref{tab:ML1_likelihoods}). Respective results are summarized in Figure \ref{fig:ML1}.\\
		Considering the changes in information entropy, it can be seen that the uncertainty about layer interface positions in $z$ was clearly reduced in the footwall. Information entropy was also decreased in the middle section of the hanging wall, but remained high in upper and lower parts, presumably due to the remaining influence of the offset uncertainty.\\
		Regarding trap volume control (Figure \ref{fig:ML1} (c)), only spill point-related migration ("SPILL POINT") and cross-fault leakage due to juxtaposition with permeable layers below the seal in the hanging wall ("LEAK UNDER") remain relevant in the posterior model. As no occurrences of leakage through overlying layers was observed, the two other mechanisms increased in frequency, while staying proportionally the same to one another. It can be presumed, that by introducing the respective likelihoods, a high thickness of the seal was ensured, as can also be recognized in Figure \ref{fig:ML1} (b). This not only made stratigraphic leakage in this model impossible, but also enforced a favorable $SSF$. A minor emergence of unrecognized control factors ("UNCLEAR") suggests that, for a few iterations, the model was possibly not correctly computable or the structural analysis failed for a different reason.\\
		The bulk that accounted for the positive mode in the prior $ROV$ distribution was raised and shifted laterally to higher values, while the probability of failure ($ROV = 0$) was diminished significantly (see Figure Figure \ref{fig:ML1} (d)). This can presumably be explained by the disappearing of leakage to overlying formations as a trap control mechanism, which was programmed to always lead to complete failure of the trap volume.\\
		As the probability of failure and thus the bimodality of the probability distribution disappeared, all Bayes actions were shifted to higher estimates, most significantly those of the two risk-averse individuals (see Figure \ref{fig:ML1} (e). These have thus converged into an area of similar expectations. Additionally, expected losses were lowered overall. The posterior state of knowledge, provided by the updated model, encouraged even the most risk-averse decision-maker to expect recoverable oil volumes from 3.5 million m$^3$ upwards.

		\subsection{Posterior model II: Likely thin reservoir formation}%ML3
		For the second model, thickness likelihoods were defined in a way that reinforced the probability of a very thick seal but at the same time restricted the reservoir formation to a likely significantly thinner thickness of around 50 m (see Table in Appendix). The results (see Appendix) are very similar, in fact practically identical, to those of model I above. This can be explained by the influence of leak and spill point. As the shape of the general structure itself was not defined to be uncertain, the position of the anticlinal spill point relative to the fault top remains mostly unchanged, irrespective of layer thickness variations. Thus, reservoir thickness is only a secondary factor to trap volume control. Assuming fixed values for all other parameters, the maximum fill horizon is most likely the same for a 300 m as for a 50 m thick reservoir unit. The thickness of the reservoir only becomes significant, if it is thin enough to locate the respective layer bottom above the leak and spill point depth in the trap section. Ultimately, the spill point appears to be the primariy limiting factor to maximum trap fill. It can furthermore be derived from this that the maximum trap volume potential, as defined by the given structural features, is already represented in the $ROV$ probability distribution of the prior model. Examining this in Figure \ref{fig:ML1} (d), it seems that this overall maximum volume is found at approximately 5.7 millon m$3$. Higher values were not observed in any other model realization and are presumed to be unattainable due to the restriction given by the spill point.

		\subsection{Posterior model III: Likely very thin seal}%ML2
		\begin{figure}[p!]
			\centering
			\includegraphics[width=1\textwidth]{Figures/ML2}
			\caption{....}\label{fig:ML2}
		\end{figure}
		Results from this third model show that in contrast to the reservoir thickness, seal thickness is a primary and decisive factor for the manifestation of a reliable trap structure. Using likelihoods which favor a thick reservoir, but only a thin seal of about 50 m thickness, leakage due to juxtaposition with seal-overlying formations across the fault ("LEAK OVER") becomes the primary mechanism for trap volume control (see Figure \ref{fig:ML2} (c). It can be recognized in Figure \ref{fig:ML2} (b) that, while uncertainty was reduced clearly in the footwall, it remains high in the hanging wall. The state of information entropy furthermore suggests that the seal layer frequently appeared significantly deeper in the hanging wall than in the footwall. Combined with a generally very thin seal, this allowed for the occurrence of high displacement to seal thickness ratios, i.e. a high $SSF$. Thus, frequent surpassing of the critical $SSF$ and subsequent fault seal breaching were enabled. It is additionally apparent in Figure \ref{fig:ML2} (b), that while fault-related leakage has become predominant, control by the spill point was suppressed to be widely irrelevant in this posterior model.\\	%It is suggested by the footwall information entropy in Figure \ref{label} (b), that the general shape of the structure has been flattened.
		A dramatically higher probability of complete trap failure is induced by the high frequency of cross-fault leakage to overlying formations. Consequently, the bulk of the posterior $ROV$ probability distribution is shifted to a volume of zero, as depicted in Figure \ref{fig:ML2} (d). Any positive trap volumes have become highly improbable.\\
		Bayes actions changed accordingly. Given this posterior $ROV$ distribution, all actors find their minimum of expected loss at $ROV = 0$, which equals the decision to "take no action" (see Figure \ref{fig:ML2} (e)). It has to be pointed out that all Bayes actions are found at the same point due to the lack of negative values into which more risk-averse actors would presumably move, given the possibility. This is a difference to the score system implemented for the 1D geological model evaluation. Volumes cannot be negative, thus, Bayes actions cannot be shifted past the limit of $ROV = 0$. Compared to the prior model, expected losses were decreased overall, but are higher for more risk-friendly individuals at the minimum.

		\subsection{Posterior model IV: SSF likelihood reinforcing trap reliability}%ML4
		For this fourth model, thickness likelihoods were chosen to be comparable to those of model I, i.e. supporting the probability of thick layers in general. Additionally, a $SSF$-related likelihood function was implemented. With a mean of 1.5 and a standard deviation of 0.5, it was chosen to mostly reinforce the likelihood of a $SSF$ below the critical value of $SSF_c = 3$.\\
		In Figure \ref{ML4} (b), it can be seen that the uncertainty about the position of interfaces in the footwall was greatly reduced for all layers. In the hanging wall, information entropies remained high but were shifted upwards and mainly concentrated in an area that is juxtaposed to the footwall seal formation.\\
		In the resulting $ROV$ distribution, the possibility of trap failure was completely extinguished, while the probability of high volumes in the range between 3.5  and 5 million m$^3$ was significantly increased. Mean and median are found at almost the same value, but the bulk of the distribution found in this range appears to be slightly bimodal in itself, resembling approximately the shape of two normal distributions overlapping around a value of 4 million m$m^3$.\\
		Accordingly, Bayes actions converged within this range of positive values. Considering the two modes in this range, risk-averse to risk-neutral bids remained in the middle of the mode of lower volumes, while risk-friendly actors were encouraged to estimate higher volumes as found in the second mode. Overall, significant convergence was achieved, also in respect to expected losses.\\		
		The importance of a reliable seal is once again emphasized by these results. Reliable fault sealing was favored by the chosen $SSF$ likelihood. It should be pointed out that computation of the complete model is required to attain the $SSF$. This value is furthermore characterized by a complex interdependence with other parameters and is most directly related to layer thicknesses and fault offset. For this reason, information entropy and offset uncertainty in the hanging wall was not greatly reduced. However, the certainty about the ratio of displacement to layer thicknesses was improved by using the $SSF$ in a likelihood function.

		\subsection{Posterior model V: Controversial information about the Shale Smear Factor}%ML5
		In this last model simulation, no thickness likelihoods, only a $SSF$-related likelihood was used. In contrast to model IV, this was chosen to narrow the probability of the $SSF$ to be around the threshold of its critical value, by basing it on a normal distribution with a mean of 3 and a standard deviation of 0.3.\\
		Results are relatively similar to those of the prior model. It is notable that the information entropy regarding the position of layer boundaries in the footwall was decreased, while the entropy seems to have been raised significantly in the hanging wall (Figure \ref{ML5} (b)). Furthermore, an increase in the relevance of cross-fault leakage to overlying formations ("LEAK OVER") as a trap volume mechanism (increase from $\sim13\5$ to $\sim18\%$) was observed (Figure \ref{ML5} (c)). Consequently, the bimodality between the bulk of positive values and the probability of complete failure in the $ROV$ distribution was apparently reinforced. This is also indicated by a shift of mean and median to lower volumes (Figure \ref{ML5} (d)). Accordingly, a higher degree of decision divergence, i.e. a greater separation of different Bayes actions, was induced (Figure \ref{ML5} (e)). Expected losses of the minima were raised as well.\\
		It is notable that the implementation of a likelihood which can be expected to significantly decrease the $SSF$ related uncertainty, overall leads to an amplification of the duality in the $ROV$ probability distribution and to dispersion of different risk-affine decisions. This effect is presumably caused by conservation of decisive uncertainty between "success" and complete failure of the trap, by narrowing the probability of the $SSF$ to be around its critical value.
